---
title: "Multiple-choice quiz | Select"
---

```{r setup, include=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, knitr, here)
knitr::opts_knit$set(upload.fun = imgur_upload)
```

# Intro

This is your first quiz on {dplyr} ! This first quiz will explore the select verb and its theory.

For multiple choice questions, please enter the correct letter option into the provided box.
For true or false question please write in T for TRUE or F for FASLE. 

------------------------------------------------------------------------

# Q: 1

Is the following statement true or false?

*Select allows to keep or drop rows.* 

## A

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

## Incorrect message

select() does not act on rows.

------------------------------------------------------------------------

# Q: 2

One of these snippets does not work for selecting columns: which one? 

## A 

`select(c(age, weight, height))`

## B

`select(age,weight,height)`

## C

`select("age", "weight", "height")`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 3

How do you select a range of columns? {c}

## A 

`select(sample1;sample10)`

## B

`select(sample1-sample10)`

## C

`select(sample1:sample10)`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 4

Is the following statement true or false?

*You can select using indexing.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 5

How many symbols can you use to drop columns ?

## A 

`1`

## B

`2`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 6

Which of the following statements is INCORRECT (false) about everything() ?

## A 

`everything() can only be used alone inside of select`

## B

`everything() selects everything`

## C

`everything() can help to reorder the dataframe`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 7

You want to select columns called: "treatmentA", "treatmentB", "treatementC". Which of these helper verbs would NOT help you in your selection ? 

## A 

`contains()`

## B

`ends_with()`

## C

`starts_with()`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 8

How do you change the naming using rename()?

## A 

`rename(OLDNAME <- NEWNAME)`

## B

`rename(OLDNAME = NEWNAME)`

## C

`rename(NEWNAME = OLDNAME)`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------
