---
title: "Multiple-choice quiz | Conditional Mutate"
---

```{r setup, include=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, knitr, here)
knitr::opts_knit$set(upload.fun = imgur_upload)
```

# Intro

We are continuing our quizzes on different {dplyr} verbs ! This quiz will check your understanding of the case_when helper function, used inside of the mutate verb.

------------------------------------------------------------------------

# Q: 1

Which of the following explains what case_when() does?

## A 

`It allows to do other operations like select, filter inside of mutate`

## B

`It allows global transformation/creation of a variable`

## C

`It allows conditional transformation/creation of a variable`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 2

Is the following statement true or false?

*case_when() can be used outside mutate () or in other {dplyr} verbs.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 3

Which of the following statements is correct about the TRUE default argument?

## A 

`It only applies to boolean variables`

## B

`It matches any row not yet matched`

## C

`It overrides all other previous conditions`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 4

Is the following statement true or false?

*The order of the conditions in case_when() does not matter.*

## A 

`It only applies to boolean variables`

## B

`It matches any row not yet matched`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 5

Is the following statement true or false? 

*The TRUE default argument should always be defined first in the case_when statement.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 6

Is the following statement true or false? 

*The TRUE default argument is the best way to handle NA values in a case_when statement.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 7

Is the following statement true or false? 

*The TRUE default argument can be used to keep the default values of a variable.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 8

What does the following pseudo-code do? 

`mutate(variable = case_when(condition1 ~ transformation1,`
                            `condition2 ~ transformation2,`
                            `TRUE ~ variable))`
                          
## A 

`It creates a new variable if the conditions are true`

## B

`It keeps the default values of the variable if the conditions are not matched`

## C

`It sets all values to the same default value if the conditions are not matched`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 9

Is the following statement true or false? 

*The LHS (left hand side) of the case_when conditions can only have one condition*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 10

Which of the following statements is correct about the order of priority of conditions in case_when ?

## A 

`The first condition is applied first, others are applied sequentially in a descending order`

## B

`The last condition is applied first, and others are applied sequentially in an ascending order`

## C

`The conditions are independent and override each other`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 11

Is the following statement true or false? 

*The order of priority of conditions in case_when matters the most when the conditions are open-ended.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 12

How can you make best usage of the order of priority of conditions in case_when statements?

## A 

`Never use open-ended conditions`

## B

`Use conditions with only one logical operator`

## C

`Use closed conditions`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 13

Is the following statement true or false? 

*if_else statements are made to be used inside case_when statements.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 14

Which of the following is correct about if_else?  

## A 

`if_else uses a tilde (~) like case_when`

## B

`if_else can only handle binary conditions`

## C

`if_else can be used outside of mutate`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------
