---
title: "Multiple-choice quiz | Filter"
---

```{r setup, include=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, knitr, here)
knitr::opts_knit$set(upload.fun = imgur_upload)
```

# Intro

We are continuing our quizzes on different {dplyr} verbs ! This quiz will check your understanding of the filter verb.

------------------------------------------------------------------------

# Q: 1

Is the following statement true or false?

*Filter allows to keep or drop rows.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 2

Which of the following is NOT a relational operator? 

## A 

`=`

## B

`<`

## C

`/`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 3

In a variable representing countries, how do you filter to only keep Swiss-relevant information? 

## A 

`filter(country==Switzerland)`

## B

`filter(country<-Switzerland)`

## C

`filter(country=Switzerland)`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

## Q: 4

What is the meaning of `A %in% B`?

## A 

`B is an element of A`

## B

`A is an element of B`

## C

`A and B are distinct`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 5

What is the following filter statement doing?

*filter(age<20 & (obese==TRUE | underweight==TRUE))*

## A 

`Keeping only old people`

## B

`Keeping people with a normal body weight`

## C

`Keeping people with an abnormal body weight`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 6

What is the following filter statement doing?

*filter(!(sex==M | age<13))*

## A 

`Keeping women older than 13`

## B

`Keeping men and women but only those older than 13`

## C

`Keeping all women but only men older than 13`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 7

How do you filter for NA values ? 

## A 

`filter(id=="NA")`

## B

`filter(id==NA)`

## C

`filter(is.na(id))`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 8

Which of the following is NOT a correct way of feeding the conditions into filter ?

## A 

`filter(condition1 & condition2 | condition3)`

## B

`filter(condition1, condition2, condition3)`

## C

`filter(condition1 + condition2 | condition3)`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------
