---
title: "Multiple-choice quiz | Mutate"
---

```{r setup, include=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, knitr, here)
knitr::opts_knit$set(upload.fun = imgur_upload)
```

# Intro

We are continuing our quizzes on different {dplyr} verbs ! This quiz will check your understanding of the mutate verb.

------------------------------------------------------------------------

# Q: 1

How do you write a mutate statement? 

## A 

`mutate(new_column = operation)`

## B

`mutate(new_column == operation)`

## C

`mutate(operation <- new_column)`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 2

How do you modify a variable in-place? 

## A 

`By adding "inplace=TRUE" inside mutate`

## B

`By keeping the same column name`

## C

`You cannot modify in-place with mutate`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 3

How do you create a rank variable using mutate?

## A 

`mutate(rank(age, "min"))`

## B

`mutate(age_rank = rank("min", age))`

## C

`mutate(rank_age = rank(age, ties.method = "min"))`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 4

What type of variable does the following line of code generate?

*mutate(male = sex==Male)*

## A 

`numeric`

## B

`boolean`

## C

`string`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 5

What does the janitor::tably() function do?

## A 

`It cleans the column names`

## B

`It makes a table with the observation counts`

## C

`It deletes NAs`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 6

Is this statement true or false?

*as.factor() can only be used inside of mutate()*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------
