---
title: "Multiple-choice quiz | Groupby Summarize"
---

```{r setup, include=FALSE}
# Load packages 
if(!require(pacman)) install.packages("pacman")
pacman::p_load(tidyverse, knitr, here)
knitr::opts_knit$set(upload.fun = imgur_upload)
```

# Intro

We are continuing our quizzes on different {dplyr} verbs ! This quiz will check your understanding of the summarize and the group_by verbs.

------------------------------------------------------------------------

# Q: 1

Which of the following defines a summary statistic?

## A 

`a summary statistic describes a value with a sequence of values`

## B

`a summary statistic describes a sequence of values with a mean`

## C

`a summary statistic describes a sequence of values with a single value`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 2

Which of the following is not a summary statistic? 

## A 

`the mean`

## B

`the division`

## C

`the inter-quartile ratio`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 3

Is the following statement true or false?

*summarize() is the only way to get the mean of a variable.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 4

How do you write a summarize statement to get the mean of the age variable ?

## A 

`summarize(mean_age = mean(age))`

## B

`summarize(mean(age))`

## C

`summarize(mean_age = mean(mean_age$age))`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 5

Is the following statement true or false?

*With summarize, you need to calculate one summary statistic at a time.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 6

Is the following statement true or false?

*The sum(), mean(), and other summary functions, automatically exclude NA values from their calculations.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 7

Is the following statement true or false?

*In group_by, the operations are split according to these groups.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 8

What does grouping by both age and sex imply  ?  

## A 

`we group by age, then we group by sex`

## B

`we generate all the unique combinations from both variables`

## C

`we create groups only for the respondents who have a recorded age or sex`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 9

The .drop argument of group_by() is important because...  ?

## A 

`It serves to show all combinations of variables, even those not accounted for in the data`

## B

`It discards the variables used to make the groupings and only keeps the groups`

## C

`It drops the groups with NA values`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 10

Fill in the blank

*After you group, using group_by(), and transforming your data according to the groups, you need to remember to ...*

## A 

`ungroup()`

## B 

`no_group()`

## C

`undo_group()`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 11

Is the following statement true or false?

*When you group_by() then summarize(), you summarize by group.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 12

Is the following statement true or false?

*When you group_by() then summarize(), you get a different summary statistic for each group.*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 13

Fill in the blank

*Grouping by age, sex, and education, all at the same time, is called ... grouping.*


## A 

`complex`

## B 

`embedded`

## C

`nested`

## Answer

C

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 14

Is the following statement true or false?

*When grouping by more than one variable, the order of the variables matters because, for example, a different order will generate different summary statistics.*

## A 

`False`

## B

`True`

## Answer

A

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------

# Q: 15

Is the following statement true or false?

*Using n() inside of summarize() is the same as using count()*

## A 

`False`

## B

`True`

## Answer

B

## Correct message

```{r echo = FALSE, results = "asis"}
cat(praise::praise())
```

------------------------------------------------------------------------
